---
title: "Exploratory visualizations"
author: "Joe Brew and Alberto García-Basteiro"
date: "January 20, 2017"
fig_height: 5
fig_width: 7
output: ioslides_presentation

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r, results = 'hide'}
# Packages
library(ggplot2)
library(cism)
library(rworldmap)
library(rgeos)
library(maptools)
library(rgdal)
library(tidyr)
library(RColorBrewer)

# Read in data
source('make_wide.R')

# Get world map
world <- readOGR('shp/world', 'ne_110m_admin_0_countries')

```

# Charts

joe@economicsofmalaria.com

## Overall correlation

```{r}
ggplot(data = df,
       aes(x = w_both_all_tbtotal_nd,
           y = i_both_all_tbtotal_nd)) +
  geom_point(color = 'darkgreen', alpha = 0.6) +
  theme_cism() +
  labs(x = 'WHO',
       y = 'IHME',
       title = 'Correlation between WHO and IHME mortality numbers',
       subtitle = 'Log scale, includes both TB and TB+HIV cases') +
  geom_abline(color = 'darkorange',
              alpha = 0.6) +
    scale_y_log10(limits = c(1, 150000)) +
  scale_x_log10(limits = c(1, 150000)) 
```

## Region-specific breakdown

The below scatterplot shows the correlation between WHO (x-axis) estimates and IHME (y-axis) estimates, with each point colored by its (WHO-defined) region. 
```{r}
cols <- colorRampPalette(brewer.pal(n = 9, name = 'Spectral'))(length(unique(df$who_region)))
ggplot(data = df,
       aes(x = w_both_all_tbtotal_nd,
           y = i_both_all_tbtotal_nd,
           color = who_region)) +
  geom_point(alpha = 0.6) +
  theme_cism() +
    scale_y_log10(limits = c(1, 150000)) +
  scale_x_log10(limits = c(1, 150000)) +
  labs(x = 'WHO',
       y = 'IHME',
       title = 'Correlation between WHO and IHME mortality numbers',
       subtitle = 'Log scale, includes both TB and TB+HIV cases') +
   # geom_smooth(method='lm', se = FALSE, alpha = 0.6) +
  geom_abline(color = 'darkorange',
              alpha = 0.6) +
  scale_color_manual(name = 'Region',
                     values = cols)
```

## Region-specific breakdown 

We can also fit each region's line of best fit. 


```{r}
cols <- colorRampPalette(brewer.pal(n = 9, name = 'Spectral'))(length(unique(df$who_region)))
ggplot(data = df,
       aes(x = w_both_all_tbtotal_nd,
           y = i_both_all_tbtotal_nd,
           color = who_region)) +
  geom_point(alpha = 0.6) +
  theme_cism() +
    scale_y_log10(limits = c(1, 150000)) +
  scale_x_log10(limits = c(1, 150000)) +
  labs(x = 'WHO',
       y = 'IHME',
       title = 'Correlation between WHO and IHME mortality numbers',
       subtitle = 'Log scale, includes both TB and TB+HIV cases') +
   geom_smooth(method='lm', se = FALSE, alpha = 0.2,
               formula = y ~ x) +
  # geom_abline(color = 'darkorange',
  #             alpha = 0.6) +
  scale_color_manual(name = 'Region',
                     values = cols)
```


## Region-specific breakdown 


We can also examine relative difference in estimates, by WHO region. TIn the below violin charts, the jittered points represent each country.

```{r}
temp <- df %>%
  filter(!is.na(who_region))
cols <- colorRampPalette(brewer.pal(n = 9, name = 'Spectral'))(length(unique(df$who_region)))

  
ggplot(data = temp,
       aes(x = who_region,
           y = relative_difference)) +
  geom_jitter(aes(color = who_region), #color = 'darkgreen', 
              alpha = 0.8) +
  geom_violin(aes(fill = who_region),
    # fill = 'darkorange', 
    alpha = 0.3) +
  theme_cism() +
  labs(x = 'WHO Region',
       y = 'Relative difference',
       title = 'Variance in estimates between WHO and IHME',
       subtitle = 'Absolute difference divided by maximum value') +
  scale_color_manual(name = '',
                     values = cols) +
  scale_fill_manual(name = '',
                    values = cols) +
  guides(color = FALSE,
         fill = FALSE)
```


## Region-specific breakdown 


The above metric of relative variance is _independent_ of which source is higher. Here, we visualize variance by examining the ratio of IHME deaths over WHO deaths.

```{r}
temp <- df %>%
  filter(!is.na(who_region))
cols <- colorRampPalette(brewer.pal(n = 9, name = 'Spectral'))(length(unique(df$who_region)))

g <- 
  ggplot(data = temp,
       aes(x = who_region,
           y = i_over_w)) +
  geom_jitter(aes(color = who_region), #color = 'darkgreen', 
              alpha = 0.8) +
  geom_violin(aes(fill = who_region),
    # fill = 'darkorange', 
    alpha = 0.3) +
  theme_cism() +
  labs(x = 'WHO Region',
       y = 'Relative difference',
       title = 'TB+TB/HIV deaths: IHME divided by WHO',
       subtitle = '') +
  scale_color_manual(name = '',
                     values = cols) +
  scale_fill_manual(name = '',
                    values = cols) +
  guides(color = FALSE,
         fill = FALSE)
print(g)
```


## Region-specific breakdown 


For ease of interpretation, we remove those outliers.

```{r}
g <- g +
  ylim(0,5)
g
```


## Region-specific breakdown 

As a complement to the above, we can also examine the ratio of WHO deaths over IHME deaths.

```{r}

temp <- df %>%
  filter(!is.na(who_region))
cols <- colorRampPalette(brewer.pal(n = 9, name = 'Spectral'))(length(unique(df$who_region)))

g <- 
  ggplot(data = temp,
       aes(x = who_region,
           y = w_over_i)) +
  geom_jitter(aes(color = who_region), #color = 'darkgreen', 
              alpha = 0.8) +
  geom_violin(aes(fill = who_region),
    # fill = 'darkorange', 
    alpha = 0.3) +
  theme_cism() +
  labs(x = 'WHO Region',
       y = 'Relative difference',
       title = 'TB+TB/HIV deaths: WHO divided by IHME',
       subtitle = '') +
  scale_color_manual(name = '',
                     values = cols) +
  scale_fill_manual(name = '',
                    values = cols) +
  guides(color = FALSE,
         fill = FALSE)
print(g)
```

# By country

## By country

In the below chart we identify those countries for whom the ratio of IHME-estimated deaths (TB + HIV/TB) is greatest (top 10) and least (bottom 10). 

```{r}
temp <- df
temp <- temp %>% filter(!is.na(i_over_w)) %>% arrange(i_over_w)
temp <- temp[c(1:10,
               (nrow(temp) - 9):nrow(temp)),]
temp$country <- factor(temp$country,
                       levels = temp$country)
ggplot(data = temp,
       aes(x = country,
           y = i_over_w)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  theme_cism() +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab('Ratio') +
  xlab('Country') +
  ggtitle('IHME over WHO: top and bottom 10')
```


## By country

As a complement to the above, we can examine the inverse: WHO deaths over IHME deaths.


```{r}
temp <- df
temp <- temp %>% filter(!is.na(w_over_i)) %>% arrange(w_over_i)
temp <- temp[c(1:10,
               (nrow(temp) - 9):nrow(temp)),]
temp$country <- factor(temp$country,
                       levels = temp$country)
ggplot(data = temp,
       aes(x = country,
           y = w_over_i)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  theme_cism() +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab('Ratio') +
  xlab('Country') +
  ggtitle('WHO over IHME: top and bottom 10')
```


## By country

Rather than ratios, we can also examine absolute differences. The below bar chart shows those countries with the greatest absolute difference (IHME estimated deaths minus WHO estimated deaths).

```{r}
temp <- df
temp <- temp %>% filter(!is.na(i_minus_w)) %>% arrange(i_minus_w)
temp <- temp[c(1:10,
               (nrow(temp) - 9):nrow(temp)),]
temp$country <- factor(temp$country,
                       levels = temp$country)
ggplot(data = temp,
       aes(x = country,
           y = i_minus_w)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  theme_cism() +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab('Difference in deaths') +
  xlab('Country') +
  ggtitle('IHME minus WHO: top and bottom 10') +
  geom_abline(yintercept = 0, alpha = 0.6, color = 'red')

```


## By country

As a complement to the above, we can do WHO deaths minus IHME deaths.

```{r}
temp <- df
temp <- temp %>% filter(!is.na(w_minus_i)) %>% arrange(w_minus_i)
temp <- temp[c(1:10,
               (nrow(temp) - 9):nrow(temp)),]
temp$country <- factor(temp$country,
                       levels = temp$country)
ggplot(data = temp,
       aes(x = country,
           y = w_minus_i)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  theme_cism() +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab('Difference in deaths') +
  xlab('Country') +
  ggtitle('WHO minus IHME: top and bottom 10') +
  geom_abline(yintercept = 0, alpha = 0.6, color = 'red')
```


## By country

All of the above are biased by small numbers (ie, countries with very few cases are most likely to have the most extreme ratios). Accordingly, we can instead examine an indicator suggested by Frank: The number of reported cases, divided by deaths.

This indicator is, in fact, two indicators. One is the reported cases divided by deaths (per the WHO), and the other is per the IHME. The below histograms show the country-specific distribution of these indicators.


## By country

Here is the distribution of the cases over deaths indicators.

```{r}
g1 <- 
  ggplot(data = df,
       aes(x = w_cases_over_deaths)) +
  geom_histogram(color = 'darkgreen', 
                 fill = 'darkorange',
                 alpha = 0.6) +
  theme_cism() +
  xlab('Cases over deaths') +
  ylab('Count') +
  ggtitle('Cases over deaths (WHO)')
g2 <- 
  ggplot(data = df,
       aes(x = i_cases_over_deaths)) +
  geom_histogram(color = 'darkgreen', 
                 fill = 'darkorange',
                 alpha = 0.6) +
  theme_cism() +
  xlab('Cases over deaths') +
  ylab('Count') +
  ggtitle('Cases over deaths (IHME)')
Rmisc::multiplot(g1, g2)
```



## By country

When we examine the cases over deaths metrics at the country specific level as well. As per before, we'll take the top and bottom 10 countries only.

```{r}
temp <- df
temp <- temp %>% filter(!is.na(w_cases_over_deaths)) %>% arrange(w_cases_over_deaths)
temp <- temp[c(1:10,
               (nrow(temp) - 9):nrow(temp)),]
temp$country <- factor(temp$country,
                       levels = temp$country)
ggplot(data = temp,
       aes(x = country,
           y = w_cases_over_deaths)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  theme_cism() +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab('Ratio') +
  xlab('Country') +
  ggtitle('Cases per WHO deaths: top and bottom 10') 
```


## By country

```{r}
temp <- df
temp <- temp %>% filter(!is.na(i_cases_over_deaths)) %>% arrange(i_cases_over_deaths)
temp <- temp[c(1:10,
               (nrow(temp) - 9):nrow(temp)),]
temp$country <- factor(temp$country,
                       levels = temp$country)
ggplot(data = temp,
       aes(x = country,
           y = i_cases_over_deaths)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  theme_cism() +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab('Ratio') +
  xlab('Country') +
  ggtitle('Cases per IHME deaths: top and bottom 10') 
```

# Rifampicin resistance

## Rifampicin resistant TB 

The below chart shows the estimated percentage of new TB cases with rifampicin resistant TB (x-axis) and the cases per WHO deaths (y-axis):

```{r}
ggplot(data = df,
       aes(x = e_rr_pct_new,
          y = w_cases_over_deaths)) +
  geom_point(color = 'darkorange') +
  theme_cism() +
  labs(x = 'Percentage rifampicin resistant',
       y = 'Ratio of cases over deaths',
       title = 'Resistance and cases over deaths',
       subtitle = 'WHO')
```


## Rifampicin resistant TB 

The below chart is identical, but using IHME deaths rather than WHO.


```{r}
ggplot(data = df,
       aes(x = e_rr_pct_new,
          y = i_cases_over_deaths)) +
  geom_point(color = 'darkorange') +
  theme_cism() +
  labs(x = 'Percentage rifampicin resistant',
       y = 'Ratio of cases over deaths',
       title = 'Resistance and cases over deaths',
       subtitle = 'IHME')
```


## Rifampicin resistant TB 

We can also examine the same x-axis (resistance), but relative to other indicators.

```{r}
ggplot(data = df,
       aes(x = e_rr_pct_new,
          y = i_over_w)) +
  geom_point(color = 'darkorange') +
  theme_cism() +
  labs(x = 'Percentage rifampicin resistant',
       y = 'Ratio: IHME over WHO deaths',
       title = 'Resistance and differential deaths',
       subtitle = 'IHME over WHO')

```

## Rifampicin resistant TB 

```{r}
ggplot(data = df,
       aes(x = e_rr_pct_new,
          y = w_over_i)) +
  geom_point(color = 'darkorange') +
  theme_cism() +
  labs(x = 'Percentage rifampicin resistant',
       y = 'Ratio: WHO over IHME deaths',
       title = 'Resistance and differential deaths',
       subtitle = 'WHO over IHME')
```

## Rifampicin resistant TB 


```{r}
ggplot(data = df,
       aes(x = e_rr_pct_new,
          y = relative_difference)) +
  geom_point(color = 'darkorange') +
  theme_cism() +
  labs(x = 'Percentage rifampicin resistant',
       y = 'Ratio',
       title = 'Resistance and differential deaths',
       subtitle = 'Relative difference: absolute difference over maximum of two estimates')
```

## Rifampicin resistant TB 


```{r}
ggplot(data = df,
       aes(x = e_rr_pct_new,
          y = absolute_difference)) +
  geom_point(color = 'darkorange') +
  theme_cism() +
  labs(x = 'Percentage rifampicin resistant',
       y = 'Deaths',
       title = 'Resistance and differential deaths',
       subtitle = 'Absolute difference in WHO / IHME estimates')
```

# Maps

## Comparative choropleths

```{r}
# Expand
world_tidy <- broom::tidy(world, region = 'iso_a3')
# Join in data
world@data <-
  left_join(x = world@data,
            y = df,
            by = c('iso_a3' = 'iso3'))
# Join to the tidy version too
world_tidy <-
  left_join(x = world_tidy,
            y = df,
            by = c('id' = 'iso3'))

# Remove those which are unknown
world_tidy <- world_tidy %>%
  filter(id != '-99',
         !is.na(country))

# Plot
g1 <- 
  ggplot(data = world_tidy,
       aes(x = long,
           y = lat,
           group = group,
           fill = i_both_all_tbtotal_nd)) +
  coord_map() +
  geom_polygon(color = 'darkorange',  size = 0.1) +
  theme_cism() +
  scale_fill_gradient2(#low = 'darkgreen',
                       #mid = 'white',
                      #   high = 'darkorange',
                       name = '') +
  labs(x = 'Longitude',
       y = 'Latitude',
       title = 'IHME',
       subtitle = 'Number of deaths (TB + TB/HIV)')
g2 <- 
  ggplot(data = world_tidy,
       aes(x = long,
           y = lat,
           group = group,
           fill = w_both_all_tbtotal_nd)) +
  coord_map() +
  geom_polygon(color = 'darkorange',  size = 0.1) +
  theme_cism() +
  scale_fill_gradient2(#low = 'darkgreen',
                       #mid = 'white',
                      #   high = 'darkorange',
                       name = '') +
  labs(x = 'Longitude',
       y = 'Latitude',
       title = 'WHO',
       subtitle = 'Number of deaths (TB + TB/HIV)') 


```

### IHME

```{r}
g1
```

### WHO

```{r}
g2
```

## Choropleth of absolute differences

```{r}
ggplot(data = world_tidy,
       aes(x = long,
           y = lat,
           group = group,
           fill = absolute_difference)) +
  coord_map() +
  geom_polygon(color = 'darkorange',  size = 0.1) +
  theme_cism() +
  scale_fill_gradient2(#low = 'darkgreen',
                       #mid = 'white',
                      #   high = 'darkorange',
                       name = '') +
  labs(x = 'Longitude',
       y = 'Latitude',
       title = 'Absolute difference',
       subtitle = 'Number of deaths (TB + TB/HIV)') 
```


## Choropleth of relative differences

```{r}
ggplot(data = world_tidy,
       aes(x = long,
           y = lat,
           group = group,
           fill = relative_difference)) +
  coord_map() +
  geom_polygon(color = 'darkorange',  size = 0.1) +
  theme_cism() +
  scale_fill_gradient2(#low = 'darkgreen',
                       #mid = 'white',
                      #   high = 'darkorange',
                       name = '') +
  labs(x = 'Longitude',
       y = 'Latitude',
       title = 'Relative difference',
       subtitle = 'Number of deaths (TB + TB/HIV)') 
```

## Choropleth of cases over deaths (WHO)

```{r}
ggplot(data = world_tidy,
       aes(x = long,
           y = lat,
           group = group,
           fill = w_cases_over_deaths)) +
  coord_map() +
  geom_polygon(color = 'darkorange',  size = 0.1) +
  theme_cism() +
  scale_fill_gradient2(#low = 'darkgreen',
                       #mid = 'white',
                      #   high = 'darkorange',
                       name = '') +
  labs(x = 'Longitude',
       y = 'Latitude',
       title = 'Cases per WHO deaths',
       subtitle = 'Ratio') 
```

## Choropleth of cases over deaths (WHO)

```{r}
ggplot(data = world_tidy,
       aes(x = long,
           y = lat,
           group = group,
           fill = i_cases_over_deaths)) +
  coord_map() +
  geom_polygon(color = 'darkorange',  size = 0.1) +
  theme_cism() +
  scale_fill_gradient2(#low = 'darkgreen',
                       #mid = 'white',
                      #   high = 'darkorange',
                       name = '') +
  labs(x = 'Longitude',
       y = 'Latitude',
       title = 'Cases per IHME deaths',
       subtitle = 'Ratio') 

```

# Interactive maps

## Relative difference (function of maximum estimate)

```{r}
library(leaflet)
qpal <- colorQuantile("YlGn", world@data$relative_difference, n = 5, na.color = "#bdbdbd")

popup <- paste0("<strong>Country: </strong>",
                    world@data$country,
                    "<br><strong>WHO: </strong>",
                    round(world@data$w_both_all_tbtotal_nd, digits = 2),
                    "<br><strong>IHME: </strong>",
                    round(world@data$i_both_all_tbtotal_nd, digits = 2),
                "<br><strong>Relative difference: </strong>",
                round(world@data$relative_difference, digits = 3),
                "<br><strong>WHO / IHME: </strong>",
                round(world@data$w_over_i, digits = 3),
                "<br><strong>IHME / WHO: </strong>",
                round(world@data$i_over_w, digits = 3),
                "<br><strong>Absolute difference: </strong>",
                round(world@data$absolute_difference, digits = 3),
                "<br><strong>Cases per WHO deaths </strong>",
                round(world@data$w_cases_over_deaths, digits = 3),
                "<br><strong>Cases per IHME deaths </strong>",
                round(world@data$i_cases_over_deaths, digits = 3))

leaflet(data = world) %>%
  addProviderTiles("Stamen.TonerBackground") %>% 
  setView(lng = mean(coordinates(world)[,1]),
          lat = mean(coordinates(world)[,2]),
          zoom = 2) %>%
  # clearShapes() %>% 
  # clearControls() %>% 
  addPolygons(data = world, fillColor = ~qpal(relative_difference), fillOpacity = 0.7, 
              color = "white", weight = 2, popup = popup) %>%
  addLegend(pal = qpal, values = ~relative_difference, opacity = 0.9,
            position = 'bottomright', 
            title = paste0('Relative difference', "<br>"))
```


## Relative difference (WHO over IHME)

Above 0 means that the WHO estimates n percentage higher than the IHME. Vice-versa.

```{r}
library(leaflet)
world@data$indicator <- (world@data$w_over_i * 100) - 100
qpal <- colorNumeric("Spectral", 
                     world@data$indicator, na.color = "#bdbdbd")

popup <- paste0("<strong>Country: </strong>",
                    world@data$country,
                    "<br><strong>WHO: </strong>",
                    round(world@data$w_both_all_tbtotal_nd, digits = 2),
                    "<br><strong>IHME: </strong>",
                    round(world@data$i_both_all_tbtotal_nd, digits = 2),
                "<br><strong>Relative difference: </strong>",
                round(world@data$relative_difference, digits = 3),
                "<br><strong>WHO / IHME: </strong>",
                round(world@data$w_over_i, digits = 3),
                "<br><strong>IHME / WHO: </strong>",
                round(world@data$i_over_w, digits = 3),
                "<br><strong>Absolute difference: </strong>",
                round(world@data$absolute_difference, digits = 3),
                "<br><strong>Cases per WHO deaths </strong>",
                round(world@data$w_cases_over_deaths, digits = 3),
                "<br><strong>Cases per IHME deaths </strong>",
                round(world@data$i_cases_over_deaths, digits = 3))

leaflet(data = world) %>%
  addProviderTiles("Stamen.TonerBackground") %>% 
  setView(lng = mean(coordinates(world)[,1]),
          lat = mean(coordinates(world)[,2]),
          zoom = 2) %>%
  # clearShapes() %>% 
  # clearControls() %>% 
  addPolygons(data = world, fillColor = ~qpal(indicator), fillOpacity = 0.7, 
              color = "white", weight = 2, popup = popup) %>%
  addLegend(pal = qpal, values = ~indicator, opacity = 0.9,
            position = 'bottomright', 
            title = paste0('WHO / IHME * 100 - 100', "<br>"))
```


## Relative difference (WHO over IHME)


In the previous chart, it may seem that there are no negative values. This is not the case. It's simply that when the WHO and IHME differ in the extremes, it pulls in one direction: 

```{r}
ggplot(data = world@data,
       aes(x = indicator)) +
  geom_histogram(fill = 'darkorange',
                 color = 'darkgreen',
                 alpha = 0.6) +
  theme_cism() +
  geom_vline(xintercept = 0,
             color = 'red',
             alpha = 0.6) +
  xlab('Ratio of WHO over IHME times 100 minus 100') +
  ylab('Countries') +
  ggtitle('Distribution of ratio of WHO over IHME times 100 minus 100')
```

## Absolute differences

WHO minus IHME instead of WHO divided by IHME.

```{r}

# qpal <- colorQuantile("RdYlGn", 
#                       world@data$i_minus_w, 
#                       n = 5, 
#                       na.color = "#bdbdbd")
qpal <- colorNumeric("RdYlGn", 
                      world@data$i_minus_w, 
                      na.color = "#bdbdbd")

popup <- paste0("<strong>Country: </strong>",
                    world@data$country,
                    "<br><strong>WHO: </strong>",
                    round(world@data$w_both_all_tbtotal_nd, digits = 2),
                    "<br><strong>IHME: </strong>",
                    round(world@data$i_both_all_tbtotal_nd, digits = 2),
                "<br><strong>Relative difference: </strong>",
                round(world@data$relative_difference, digits = 3),
                "<br><strong>WHO / IHME: </strong>",
                round(world@data$w_over_i, digits = 3),
                "<br><strong>IHME / WHO: </strong>",
                round(world@data$i_over_w, digits = 3),
                "<br><strong>Absolute difference: </strong>",
                round(world@data$absolute_difference, digits = 3),
                "<br><strong>Cases per WHO deaths </strong>",
                round(world@data$w_cases_over_deaths, digits = 3),
                "<br><strong>Cases per IHME deaths </strong>",
                round(world@data$i_cases_over_deaths, digits = 3))

leaflet(data = world) %>%
  addProviderTiles("Stamen.TonerBackground") %>% 
  setView(lng = mean(coordinates(world)[,1]),
          lat = mean(coordinates(world)[,2]),
          zoom = 2) %>%
  # clearShapes() %>% 
  # clearControls() %>% 
  addPolygons(data = world, 
              fillColor = ~qpal(i_minus_w), fillOpacity = 0.7, 
              color = "white", weight = 2, popup = popup) %>%
  addLegend(pal = qpal, values = ~i_minus_w, opacity = 0.9,
            position = 'bottomright', 
            title = paste0('IHME - WHO', "<br>"))
```


## Absolute differences


The previous map may appear overwhelmingly green (ie, that most of the countries have a near 0 difference). This is due to a difference in scale (ie, very few extreme observations). We can get a more granular look at things by taking the quadratic root of our indicator (next).

## Absolute differences


```{r}
# qpal <- colorQuantile("RdYlGn", 
#                       world@data$i_minus_w, 
#                       n = 5, 
#                       na.color = "#bdbdbd")
world@data$i_minus_w_sqrt <- (abs(world@data$i_minus_w)) ^ (1/4)
sign  <- ifelse(world@data$i_minus_w >= 0,
                TRUE,
                FALSE)
world@data$i_minus_w_sqrt <- 
  world@data$i_minus_w_sqrt *
  ifelse(sign, 1, -1)
qpal <- colorNumeric("RdYlGn", 
                      world@data$i_minus_w_sqrt, 
                      na.color = "#bdbdbd")

popup <- paste0("<strong>Country: </strong>",
                    world@data$country,
                    "<br><strong>WHO: </strong>",
                    round(world@data$w_both_all_tbtotal_nd, digits = 2),
                    "<br><strong>IHME: </strong>",
                    round(world@data$i_both_all_tbtotal_nd, digits = 2),
                "<br><strong>Relative difference: </strong>",
                round(world@data$relative_difference, digits = 3),
                "<br><strong>WHO / IHME: </strong>",
                round(world@data$w_over_i, digits = 3),
                "<br><strong>IHME / WHO: </strong>",
                round(world@data$i_over_w, digits = 3),
                "<br><strong>Absolute difference: </strong>",
                round(world@data$absolute_difference, digits = 3),
                "<br><strong>Cases per WHO deaths </strong>",
                round(world@data$w_cases_over_deaths, digits = 3),
                "<br><strong>Cases per IHME deaths </strong>",
                round(world@data$i_cases_over_deaths, digits = 3))

leaflet(data = world) %>%
  addProviderTiles("Stamen.TonerBackground") %>% 
  setView(lng = mean(coordinates(world)[,1]),
          lat = mean(coordinates(world)[,2]),
          zoom = 2) %>%
  # clearShapes() %>% 
  # clearControls() %>% 
  addPolygons(data = world, 
              fillColor = ~qpal(i_minus_w_sqrt), fillOpacity = 0.7, 
              color = "white", weight = 2, popup = popup) %>%
  addLegend(pal = qpal, values = ~i_minus_w_sqrt, opacity = 0.9,
            position = 'bottomright', 
            title = paste0('IHME - WHO (4 root)', "<br>"))

```